An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument

### 2016-01-12 Papers/InProgress/IJCAI/paper.tex
paper improvements

\section{Conclusion}\label{sec:conclusion}
The inconsistency of G\"odel's original axioms in his ontological
argument has been detected automatically by the prover
\textsc{Leo-II}. Here we have presented a rational reconstruction and
verification of the inconsistency proof in Isabelle/HOL. Interestingly,
the inconsistency argument applies to all normal modal logics
including K.

\begin{figure*}
\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
\end{figure*}

@@

On the other hand, there is still room for improvements in
Isabelle/HOL. For example, Isabelle/HOL's Sledgehammer tool, which
realises calls to external automated theorem provers, such as
\textsc{Leo-II}, still fails to detect the inconsistency of G\"odel's
axioms, while a direct modeling of the problem in TPTP THF syntax in
combination with a direct call of \textsc{Leo-II} succeeds. In other
words, without independent experiments directly in TPTP THF the
inconsistency would thus not have been detected. There are further
technical issues regarding Sledgehammer that motivate
improvements. For example, in default setting, Sledgehammer does not
inform the user early that a proof has been found and instead silently
applies first a series of time-consuming analysis tools (e.g. it's
minimizer), before it eventually informs the user of success. For the
involved proof problems presented here (e.g. T3), this phase of
silence may take several minutes (during which the user might actually
give up on the proof attempt), while \textsc{Leo-II} already reported
success to Sledgehammer much earlier (e.g. after 2,5 seconds for T3).

@@

@@

*

>>>>

\section{Conclusion}\label{sec:conclusion}
The inconsistency of G\"odel's original axioms in his ontological
argument has been detected automatically by the prover
\textsc{Leo-II}. Here we have presented a rational reconstruction and
verification of the inconsistency proof in Isabelle/HOL. This
inconsistency argument applies to all normal modal logics including K.
\begin{figure*}
\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
\end{figure*}

@@

On the other hand, there is still room for improvements in
Isabelle/HOL. For example, Isabelle/HOL's Sledgehammer tool, which
realises calls to external automated theorem provers, such as
\textsc{Leo-II}, still fails to detect the inconsistency of G\"odel's
axioms, while a direct modeling of the problem in TPTP THF syntax in
combination with a direct call of \textsc{Leo-II} succeeds. In other
words, without independent experiments directly in TPTP THF the
inconsistency would thus not have been detected. There are further
technical issues regarding Sledgehammer that motivate
improvements. For example, in default setting, Sledgehammer does not
inform the user early that a proof has been found and instead silently
applies first a series of time-consuming analysis tools (e.g. it's
minimizer), before it eventually informs the user of success. For the
involved proof problems presented here (e.g. G\"odel's theorem T3:
\textit{Necessarily, there exists God}), this phase of silence may
take several minutes (during which the user might actually give up on
the proof attempt), while \textsc{Leo-II} already reported success to
Sledgehammer much earlier (e.g. after 2,5 seconds for T3).

@@

Another open problem that we solved in this paper is a fully automatic
proof of T3 directly from Scott's axioms. Again, this proof was
contributed by \textsc{Leo-II}. However, this has been possible only
after providing a more efficient embedding, based on a universal
accessibility relation, for HOML S5 in HOL.

Both, the automated detection of the inconsistency in G\"odel's axioms
and the fully automatic proof of T3 from Scott's axioms demonstrates
the potential of our AI technology for philosophy: this technology is
already in its current state able to contribute novel results to
metaphysics and to conduct reasoning steps at granularity-levels
beyond common human capabilities.

@@
